### 2019-05-09 main_figures/figure1_doublets.Rmd
initial upload

---
title: "Figure 1 doublets"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true

---

```{r}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(ROCit)
  library(cowplot)
})
# we load the dataset
sce <- readRDS("../datasets/mixology10x5cl.SCE.rds")
CD <- as.data.frame(colData(sce))
```

```{r}
truth <- factor(CD$demuxlet_cls, c("SNG","DBL"))
roclist <- lapply( list(plDoubletFinder=CD$pldf.ratio,
                        DoubletFinder.p=CD$doubletFinder.pANN_0.25_0.02_82,
                        scran.score=CD$scran.doubletscore,
                        scds.hybrid=CD$hybrid_score), 
                   class=truth, FUN=rocit)
colors <- c("blue","red","black","pink")
plot(roclist[[1]], lwd=2, col=c(colors[[1]], "grey"), legend=FALSE, YIndex=FALSE, main="mixology10x5cl")
for(i in 2:length(roclist)){
  lines(roclist[[i]]$FPR, roclist[[i]]$TPR, lwd=2, col=colors[[i]])
}
legend("bottomright", fill=colors, legend=names(roclist))
```