### 2019-05-09 main_figures/figure2_filtering.Rmd

---
title: "Figure 2 filtering"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true

---