### 2019-05-09 supplementary_figures/supp_figures.Rmd

---
title: "Supplementary Figures"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: margin=1.5cm
output:
  pdf_document:
    keep_tex: true
    toc: true
    fig_width: 8
    fig_height: 6
    citation_package: natbib
bibliography: ../bmc_article.bib
---

```{r setup, include=FALSE}
# comment this out to print code
knitr::opts_chunk$set(echo=FALSE)
FIG_NB <- 0
getFigNb <- function(increment=FALSE){
  if(increment) FIG_NB <<- FIG_NB + 1
  FIG_NB
}
```

\newpage

```{r child='supp1_dataset_description.Rmd'}
```