### 2019-05-09 supplementary_figures/supp_figures.tex

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.5cm]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Supplementary Figures},
            pdfauthor={Pierre-Luc Germain},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Supplementary Figures}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Pierre-Luc Germain}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{09 Mai, 2019}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\newpage

\hypertarget{supplementary-figure-1}{%
\section{Supplementary Figure 1}\label{supplementary-figure-1}}

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-4-1.pdf}

\hypertarget{supplementary-figure-1-1}{%
\subsubsection{Supplementary Figure 1}\label{supplementary-figure-1-1}}

Distribution across cells of various control properties in the different
datasets. The lines indicate respectively 2 and 5 median absolute
deviations (MADs).

\newpage

\hypertarget{supplementary-figure-2}{%
\section{Supplementary Figure 2}\label{supplementary-figure-2}}

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-5-1.pdf}

\hypertarget{supplementary-figure-2-1}{%
\subsubsection{Supplementary Figure 2}\label{supplementary-figure-2-1}}

The total counts and total features per cell of doublets versus other
cells. We used the demuxlet annotation of doublets (based on SNPs) made
available through CellBench. While doublets tend to have a higher total
count and especially number of detected features, these features alone
are not always sufficient for their identification.

\newpage

\hypertarget{supplementary-figure-3}{%
\section{Supplementary Figure 3}\label{supplementary-figure-3}}

{[}to come{]} \#\#\# Supplementary Figure 3 Doublets that cannot be
identified by our method show lower signs of being doublets.

\newpage

\hypertarget{supplementary-figure-4}{%
\section{Supplementary Figure 4}\label{supplementary-figure-4}}

\includegraphics{supp_figures_files/figure-latex/unnamed-chunk-6-1.pdf}

\hypertarget{supplementary-figure-4-1}{%
\subsubsection{Supplementary Figure 4}\label{supplementary-figure-4-1}}

There is a tight relationship, in 10x datasets (i.e.~not the
\texttt{Koh} and \texttt{Kumar} datasets), between the total counts of a
cell and its number of detected features. We therefore include, among
control variables, deviation from this ratio.

\bibliography{../bmc\_article.bib}

\end{document}