### 2019-05-18 main_figures/figure2_filtering.Rmd
added missing datasets

---
title: "Figure 2 filtering"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true

---

```{r}
suppressPackageStartupMessages({
  library(ComplexHeatmap)
})
suppressMessages(devtools::load_all("../../pipComp/"))
source("../misc_functions.R")
data(datasets)
```

```{r figure2, fig.width=9, fig.height=4}
res.f <- readRDS("../../resNew/filtering_endSummary.rds")

res.f$pcCells <- 100-100*res.f$nbCells/datasets[as.character(res.f$dataset),"nbCells"]
res.f$method <- factor(paste(res.f$doubletmethod, res.f$filt))
res.f <- res.f[grep("clust",res.f$method, invert=TRUE),]
levels(res.f$method) <- gsub("nothingFUN ","",levels(res.f$method))

w <- which(abs(res.f$nbClusters-datasets[as.character(res.f$dataset),"subpopulations"])<=1)
#w <- which(res.f$nbClusters==datasets[as.character(res.f$dataset),"subpopulations"])
aris.1 <- cast2(res.f[intersect(w, which(res.f$norm=="norm.seurat")),], formula=method~dataset, value.var="ARI")
aris.2 <- cast2(res.f[intersect(w, which(res.f$norm=="norm.seuratvst")),], formula=method~dataset, value.var="ARI")
pc <- cast2(res.f, formula=method~dataset, value.var="pcCells")
chm(list("ARI (standard norm)"=aris.1, "ARI (sctransform)"=aris.2, "% cells out"=pc), scale=c("column","column","none"), value_format=c("%#.2f","%#.2f","%#.1f"), cluster_columns=FALSE, cluster_rows=FALSE, scaleVar=FALSE)
```