### 2019-05-18 supplementary_figures/supp2_filtering.Rmd
updated supp figures

---
title: "Supplementary Figures relative to filtering"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r, include=FALSE}
if(!exists('FIG_NB')){
  FIG_NB <- 0
  getFigNb <- function(increment=FALSE){
    if(increment) FIG_NB <<- FIG_NB + 1
    FIG_NB
  }
}
```

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(SingleCellExperiment)
  library(reshape2)
  library(ComplexHeatmap)
  library(knitr)
  library(circlize)
})
suppressMessages(devtools::load_all("../../pipComp/"))
source("../misc_functions.R")
data(datasets)
```

# Supplementary Figure `r getFigNb(TRUE)`

```{r filt_more_ari, fig.width=8, fig.height=6.5}
res <- readRDS("../data/filtering_old.rds")
res$pcCells <- round(100*res$nbCells/datasets[as.character(res$dataset),"nbCells"],2)
# we restrict to analyses that gave a number of clusters close to reality:
res2 <- res[which(abs(res$nbClusters-datasets[as.character(res$dataset), "subpopulations"])<=1),]
p1 <- ggplot(res2, aes(x=pcCells, y=ARI, colour=norm)) + geom_point() + facet_wrap(~dataset, scale="free")

fres <- readRDS("../../resNew/filtering_intermediateSummary.rds")
pc <- fres$filtering$pcCells
pc <- pc[grep("doublet",row.names(pc),invert=T),]
row.names(pc) <- reduceMethodNames(row.names(pc),2)
ha <- HeatmapAnnotation(dataset=reduceMethodNames(colnames(pc),1))
# shortDSNames <- c("mixology10x3cl"="mix3cl", "mixology10x5cl"="mix5cl", "Zhengmix4eq"="Zh4eq", "Zhengmix4uneq"="Zh4uneq", "Zhengmix8eq"="Zh8eq")
# colnames(pc) <- paste(shortDSNames[reduceMethodNames(colnames(pc),1)], reduceMethodNames(colnames(pc),2))
# p2 <- grid.grabExpr(draw(Heatmap(pc, name="% out", bottom_annotation=ha)))

pal <- colorRamp2(c(0,1,5,20,max(pc)),c("#0000FFFF", "#B38BFFFF", "#FFFFFFFF", "#FF9E81FF", "#FF0000FF"))
p2 <- grid.grabExpr(draw(Heatmap(pc, name="% out", bottom_annotation=ha, col=pal, cluster_columns=FALSE, show_column_names=FALSE)))

plot_grid(p1, p2, labels=LETTERS[1:2], nrow=2, rel_heights=c(3,2))
```

### Supplementary Figure `r getFigNb()`
**A:** Adjusted rand index (ARI) of various clustering pipelines using various combinations of filtering criteria. Only clustering analyses with a number of clusters +/-1 from the real number of clusters were considered. In general, filtering more does not result in more accurate clustering. **B:** Percentage of cells filtered out per subpopulation, using various combination of filters (see methods).

\newpage

# Supplementary Figure `r getFigNb(TRUE)`

```{r, fig.height=4}
res$filt.times <- sapply(strsplit(res$filt,"_"), FUN=function(x){ if(length(x)<3) return(1); as.numeric(x[[3]])})
res$filt.MADs <- sapply(strsplit(res$filt,"_"), FUN=function(x){ as.numeric(x[[2]]) })
filt.vars <- strsplit(sapply(strsplit(res$filt,"_"), FUN=function(x) x[[1]]),";")
fv <- unique(unlist(filt.vars))
for(f in fv[grep("both",fv,invert=TRUE)]){
  f2 <- gsub("lower$|higher$","both",f)
  res[[f]] <- sapply(filt.vars, FUN=function(x){ any(c(f,f2) %in% x) })
}
form <- paste0("~nbClusters*dataset+norm+",paste(grep(".",colnames(res),fixed=TRUE,value=TRUE),collapse="+"))
form <- gsub("filt.MADs","factor(filt.MADs)",form,fixed=T)
form <- gsub("filt.times","factor(filt.times)",form,fixed=T)
form <- paste0("ARI",form)
message(strwrap(paste("lm(", form, ")")))
mod <- lm(as.formula(form), data=res)
mod <- coef(summary(mod))[grep(".",row.names(coef(summary(mod))),fixed=T),]
colnames(mod)[4] <- "p value"
kable(mod)
```

### Supplementary Figure `r getFigNb()`
Linear regression analysis of ARI on various filtering criteria, correcting for the effect of the number of clusters. Increasing the number of distributions on which a cell must deviate in order to be excluded (`filt.times`) tended to be positively associated with ARI, while increasing the number of median absolute deviations required to be an outlier (`filt.MADs`) was associated with a decrease in ARI.

\newpage

# Supplementary Figure `r getFigNb(TRUE)`

```{r, fig.height=5}
sw <- fres$dimreduction$clust.avg.silwidth[[1]]
sw <- sw[,grep("doublet",colnames(sw))]
w <- grep("seuratvst",colnames(sw))
colnames(sw) <- reduceMethodNames(colnames(sw),2)
sw <- sw[order(reduceMethodNames(row.names(sw),1), rowMeans(sw)),]
ra <- data.frame(meanSilWidth=rowMeans(sw), dataset=reduceMethodNames(row.names(sw),1))
chm(list("standard Seurat norm"=sw[,-w], "sctransform"=sw[,w]), scale="row", scaleVar=FALSE, scaleTitle="meanSilWidth\ndifference", sameScale=TRUE, row_ann=ra, show_row_names=FALSE, cluster_rows=FALSE)
```

```{r fig.height=3.5, fig.width=6}
r <- fres$dimreduction$PC.R2[which(fres$dimreduction$PC.R2$dim<=10),]
r <- r[grep("doublet",r$method),]
r <- r[grep("clust",r$method, invert=T),]
r <- aggregate(r$value, by=r[,c(1,3)], FUN=mean)
r <- t(cast2(r, dataset~method, value.var="x", fun.aggregate=mean))
w <- grep("seuratvst", row.names(r))
row.names(r) <- reduceMethodNames(row.names(r), 2)
chm(list("standard Seurat norm"=t(r[-w,]), "sctransform"=t(r[w,])), sameScale=TRUE, 
    scale="row", scaleTitle="z-score", value_format="%.2f")
```

### Supplementary Figure `r getFigNb()`
Impact of the set of filters on the average silhouette width of the subpopulations (top) and on the proportion of variance in the first components explained by clusters (bottom).

\newpage

# Supplementary Figure `r getFigNb(TRUE)`

```{r}
res <- readRDS("../data/selection_old.rds")
res <- res[grep("deviance|seurat.variance.standardized",res$sel),]
res$filt <- as.factor(res$filt)
levels(res$filt) <- c("coding only", "no mitochondrial", "no ribosomal", "all")
res2 <- res[which(abs(res$nbClusters-datasets[res$dataset,"subpopulations"])<=1),]
ll <- lapply(split(res2, res2$norm), FUN=function(x){
  m2 <- cast2(x, formula=filt~dataset, fun.aggregate=median, value.var="ARI")
})
chm(ll, scale="column", scaleTitle="z-score", sameScale=TRUE, value_format="%.2f", row_sortFn=function(x){ median(x,na.rm=T) }, cluster_columns=FALSE)
```

### Supplementary Figure `r getFigNb()`

Impact of restricting the type of features used on the ARI of the clustering.
\newpage