### 2019-05-19 main_figures/figure4_selection.Rmd
included combined selection in panel A

---
title: "Figure 4 selection"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
})
source("../misc_functions.R")
suppressMessages(devtools::load_all("../../pipComp/"))
data(datasets)
theme_set(theme_cowplot(font_size = 12))
```

```{r}
GI <- readRDS("../data/GI.rds")
get_auc <- function(preds,ve.var="vst.R2",n=1000){
    preds$vst.Gini <- -1*preds$vst.Gini
    p2 <- as.matrix(preds[,c("seurat.variance", "seurat.variance.standardized", "seurat.res.var", "deviance")])
    preds2 <- list( deviance=preds$deviance,
                   seurat.disp=preds$seurat.dispersion,
                   seurat.disp.std=preds$seurat.dispersion.scaled,
                   seurat.var=preds$seurat.variance,
                   "seurat.var.std*"=preds$seurat.variance.standardized,
                   sctransform=preds$seurat.res.var,
                   rankMean=colMeans(colRanks(p2)),
                   rankMax=colMaxs(colRanks(p2)),
                   expression=preds$total_counts
                   )
    preds2 <- lapply(preds2, as.numeric)
    vesum <- sum(sort(preds[[ve.var]],decreasing=T)[1:n],na.rm=T)
    sapply(preds2, ve=preds[[ve.var]], FUN=function(x, ve) sum(ve[order(x, decreasing=T)[1:n]],na.rm=T))/vesum
}
# aucs1 <- melt(sapply(GI,FUN=get_auc))
# colnames(aucs1) <- c("method", "dataset", "varExplained")
# p1 <- noXlegWith(ggplot(aucs1, aes(x=method, y=varExplained, fill=method)) + geom_col())
# p1 <- p1 + ylab("Proportion of the explained variance") + facet_wrap(~dataset, scale="free",dir="v")
# plot_grid( p1, p2, labels=LETTERS[1:2], nrow=2)

aucs1 <- sapply(GI,FUN=get_auc)
aucs2 <- sapply(GI,ve.var="devianceExplained", FUN=get_auc)
p1 <- grid.grabExpr(draw(chm(list("Variance explained"=aucs1, "Deviance explained"=aucs2), sameScale=TRUE, scaleTitle="Proportion", cluster_columns=FALSE)))
```

```{r figure4, fig.width=6, fig.height=10}
res <- readRDS("../../resNew/sel_endSummary.rds")
res <- res[grep("devianceExplained|varExp", res$sel, invert=TRUE),]
res$selection <- paste(res$sel, res$selnb)
nbClusDiff <- abs(res$nbClusters-datasets[res$dataset,"subpopulations"])
#res2 <- res[which(nbClusDiff==0),]
res2 <- res[which(nbClusDiff==0 | (nbClusDiff<=1 & res$dataset=="mixology10x3cl")),]
ll <- lapply( split(res2, res2$norm), FUN=function(x){
  cast2(x, formula=selection~dataset, value.var="ARI")
})
names(ll) <- c("scran.noscale","standard norm", "sctransform")
p2 <- grid.grabExpr(draw(chm(ll[-1], sameScale=TRUE, scaleTitle="ARI", cluster_rows=FALSE, cluster_columns=FALSE)))

plot_grid(p1, p2, labels=LETTERS[1:2], nrow=2, rel_heights=c(6,10))
```

**A:** Estimating different methods of ranking genes based on their ability to capture genes with a high proportion of variance (left) or deviance (right) explained. rankMean and rankMax refer to, respectively, the mean rank and the maximum rank in the following metrics: variance, standardized variance, sctransform, deviance.
**B: ** Accuracy of clusterings (at the right number of clusters) based on the given clustering methods and number of genes selected.