### 2019-05-27 supplementary_figures/supp6_clust.Rmd

---
title: "Supplementary Figures relative to clustering"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r, include=FALSE}
if(!exists('FIG_NB')){
  FIG_NB <- 0
  getFigNb <- function(increment=FALSE){
    if(increment) FIG_NB <<- FIG_NB + 1
    FIG_NB
  }
}
```

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(reshape2)
  library(dplyr)
})
theme_set(theme_cowplot(font_size = 13))
suppressMessages(devtools::load_all("../../pipComp/"))
data(datasets)
source("../misc_functions.R")
```

# Supplementary Figure `r getFigNb(TRUE)`

```{r, fig.width=8, fig.height=8}
res <- readRDS("../data/clust_endSummary.rds")
res <- res[which(res$dims>2 & res$resolution %in% c(0.05, 0.15, 0.3, 0.5, 0.8, 1, 1.2)),]
res$method <- paste0( gsub("clust.","",res$clustmethod, fixed=T), 
                      " k=",res$k, 
                      ifelse( res$clustmethod=="clust.seurat", paste(" res=",res$resolution), paste0(" steps=",res$steps)) )
res$diff.nbClusters <- res$nbClusters-datasets[res$dataset,"subpopulations"]
ll <- lapply( unique(res$dataset), FUN=function(x){
  as.matrix(cast2(res[which(res$dataset==x),], formula=method~dims, fun.aggregate=mean, value.var="diff.nbClusters"))
})
names(ll) <- unique(res$dataset)
mmax <- ceiling(max(abs(sapply(ll,FUN=range))))
cf <- circlize::colorRamp2(c(-mmax, 0, mmax), colors=c("blue", "white", "red"))
chm(ll, sameScale=TRUE, scaleTitle="Difference\nto real nb\nsubpopulations\n", colors=cf, cluster_columns=FALSE, cluster_rows=FALSE)
```

### Supplementary Figure `r getFigNb()`
Difference between the number of detected clusters and the number of real subpopulations according to different numbers of dimensions and clustering paramters.

\newpage