### 2019-11-03 main_figures/figure3_normalization.Rmd
updated comparison of normalization methods

---
title: "Figure 3 normalization"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
})
suppressMessages(devtools::load_all("../../pipeComp/"))
source("../misc_functions.R")
theme_set(theme_cowplot(font_size = 11))
```

```{r}
res <- readRDS("../data/results_norm.rds")
```

```{r}
p1 <- scrna_evalPlot_DR(res, "covarRes", agg.by = "norm", anno_legend=FALSE) +
    scrna_evalPlot_DR(res, "covarRes", covar = "log10_total_features", agg.by = "norm", anno_legend=FALSE)
p1 <- grid.grabExpr(draw(p1))

p2 <- grid.grabExpr(draw(
  scrna_evalPlot_DR(res, what="silhouette", agg.by = "norm", value_format = "", show_heatmap_legend = TRUE)))

H <- scrna_evalPlot_clustAtTrueK(res, "ARI", agg.by = "norm", scale = TRUE, reorder_rows = TRUE, anno_legend=FALSE)
H <- H + scrna_evalPlot_clust(res, "MI", agg.by = "norm", scale = TRUE, reorder_rows=row.names(H@matrix), anno_legend=FALSE)
p3 <- grid.grabExpr(draw(H))
```

```{r figure3, fig.width=9, fig.height=9}
plot_grid(p1,p2,p3, nrow = 3)
```