### 2019-11-09 main_figures/figure4_selection.Rmd
updated selection figure

---
title: "Figure 4 selection"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true
---

```{r}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
})
source("../misc_functions.R")
suppressMessages(devtools::load_all("../../pipeComp/"))
data(datasets)
theme_set(theme_cowplot(font_size = 12))
```

```{r}
lf <- list.files("../datasets", "SCE", full.names=TRUE)
names(lf) <- gsub("\\.SCE\\.rds","",basename(lf))
GI <- lapply(lf, FUN=function(x) rowData(readRDS(x)))
get_auc <- function(preds,ve.var="vst.varExp",n=1000){
    preds$vst.Gini <- -1*preds$vst.Gini
    p2 <- as.matrix(preds[,c("seurat.variance", "seurat.variance.standardized", "seurat.res.var", "deviance")])
    preds2 <- list( deviance=preds$deviance,
                   seurat.disp=preds$seurat.dispersion,
                   seurat.disp.std=preds$seurat.dispersion.scaled,
                   seurat.var=preds$seurat.variance,
                   "seurat.var.std*"=preds$seurat.variance.standardized,
                   sctransform=preds$seurat.res.var,
                   #rankMean=colMeans(colRanks(p2)),
                   #rankMax=colMaxs(colRanks(p2)),
                   expression=preds$total_counts
                   )
    preds2 <- lapply(preds2, as.numeric)
    vesum <- sum(sort(preds[[ve.var]],decreasing=T)[1:n],na.rm=T)
    sapply(preds2, ve=preds[[ve.var]], FUN=function(x, ve) sum(ve[order(x, decreasing=T)[1:n]],na.rm=T))/vesum
}
# aucs1 <- melt(sapply(GI,FUN=get_auc))
# colnames(aucs1) <- c("method", "dataset", "varExplained")
# p1 <- noXlegWith(ggplot(aucs1, aes(x=method, y=varExplained, fill=method)) + geom_col())
# p1 <- p1 + ylab("Proportion of the explained variance") + facet_wrap(~dataset, scale="free",dir="v")
# plot_grid( p1, p2, labels=LETTERS[1:2], nrow=2)

aucs1 <- sapply(GI,FUN=get_auc)
h1 <- Heatmap(aucs1 , name="Proportion", cluster_columns=FALSE, 
           bottom_annotation=.ds_anno(colnames(aucs1),TRUE), show_heatmap_legend = TRUE,
           show_column_names = FALSE, col=viridisLite::inferno(100), column_title="Variance explained")
aucs2 <- sapply(GI,ve.var="devianceExplained", FUN=get_auc)
h2 <- Heatmap( aucs2[row.names(h1@matrix),], 
               cluster_rows=FALSE, cluster_columns=FALSE, 
              bottom_annotation=.ds_anno(colnames(aucs2),FALSE), show_heatmap_legend = FALSE,
              show_column_names = FALSE, col=viridisLite::inferno(100), column_title="Deviance explained")
p1 <- grid.grabExpr(draw(h1+h2))
```

```{r figure4, fig.width=7.5, fig.height=6.5}
res <- readRDS("../data/results_selection.rds")
cl <- res$clustering
vst <- cl[grep("seuratvst",row.names(cl)),]
snorm <- cl[grep("seuratvst",row.names(cl), invert=TRUE),]
h1 <- scrna_evalPlot_clustAtTrueK(snorm[grep("=1000;",row.names(snorm)),], "ARI", scale = FALSE, title="ARI at true # clusters", anno_legend = FALSE)
h2 <- scrna_evalPlot_clustAtTrueK(vst[grep("=1000;",row.names(vst)),], "ARI", scale = FALSE, reorder_rows=row.names(h1@matrix), title="sctransform\nARI at true # clusters", name="ARI2", anno_legend = FALSE)
p2 <- grid.grabExpr(draw(h1+h2))
plot_grid(p1, p2, labels=LETTERS[1:2], nrow=2, rel_heights = c(4,5))
```

**A:** Estimating different methods of ranking genes based on their ability to capture genes with a high proportion of variance (left) or deviance (right) explained by the subpopulations.
**B: ** Accuracy of clusterings (at the true number of clusters) when selecting 1000 genes using the given methods. Based on standard Seurat normalization (left) or sctransform (right). The methods "vst.varExp" and "devianceExplained" are the estimates used in **A** to evaluate the selection methods, and were included here only for validation purpose.