### 2019-11-27 main_figures/dr.Rmd
renamed main figures, updated filtering and dimensionality figures

---
title: "Figure 5 dimensionality reduction"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
})
source("../misc_functions.R")
suppressMessages(devtools::load_all("../../pipComp/"))
data(datasets)
theme_set(theme_cowplot(font_size = 12))
```

```{r, fig.width=8, fig.height=4}
res <- readRDS("../data/dr_intermediateSummary.rds")
sw <- res$dimreduction$clust.avg.silwidth[[1]]
sw <- sw[,grep("seuratvst",colnames(sw))]
sw <- sw[,grep("deviance",colnames(sw))]
colnames(sw) <- reduceMethodNames(colnames(sw),6)
sw <- matrix(unlist(sw),ncol=ncol(sw), dimnames=dimnames(sw))
ag <- aggregate(sw, by=list(reduceMethodNames(row.names(sw),1)), FUN=mean)
row.names(ag) <- ag[,1]; ag[,1] <- NULL
sw <- t(ag)

r2 <- res$dimreduction$PC.R2
r2 <- r2[grep("seuratvst",r2$method),]
r2 <- r2[grep("deviance",r2$method),]
r2$dr <- reduceMethodNames(r2$method, 6)
r2 <- r2[which(r2$dim <= 5),]
r3 <- cast2(r2, formula=dr~dataset, value.var="value")

res <- readRDS("../data/dr_endSummary.rds")
res2 <- res[which(abs(res$nbClusters-datasets[res$dataset,"subpopulations"])<=1 & 
                    res$norm=="norm.seuratvst"),]
m.ari <- cast2(res2, formula=dr~dataset, fun.aggregate=median, value.var="ARI")

ll <- list( "mean silhouette width"=sw,
            "PC1-5 var explained"=r3,
            "median ARI"=m.ari)
ll <- lapply(ll, FUN=function(x){
  row.names(x) <- gsub(".noW weighted", "", paste(row.names(x), "weighted"), fixed=T)
  x
})
chm(ll, scale="column", scaleTitle="z-score", sameScale=TRUE, value_format="%.2f", row_sortFn=function(x){ median(x,na.rm=T) }, cluster_columns=FALSE)
```

Mean of subpopulation average silhouette width (left), mean proportion of the variance in the first 5 components explained by real subpopulations (center), and median ARI of the resulting clustering (right).