### 2019-11-29 main_figures/dr.Rmd
updated dimensionality reduction comparison

---
title: "Figure 5 dimensionality reduction"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
})
source("../misc_functions.R")
suppressMessages(devtools::load_all("../../pipeComp/"))
```

```{r dimred, fig.width=12, fig.height=6}
res <- readRDS("../data/results_dimred.rds")
x <- res$dimreduction$clust.avg.silwidth
x <- x[grep("norm\\.seurat\\.noscale.*denoise", row.names(x), invert=TRUE),]
x <- x[order(colSums(apply(x,1,prob=c(0.05,0.5),FUN=quantile))+
                         rowMeans(x), decreasing=TRUE),order(colMeans(x), decreasing=TRUE)]
row.names(x) <- gsub("norm\\.","",apply(parsePipNames(row.names(x))[,c("norm","dr")],1,collapse=" + ",FUN=paste))
row.names(x) <- gsub("seuratvst","sctransform",row.names(x))
h1 <- Heatmap(x, cluster_columns = FALSE, cluster_rows = FALSE, 
        name="silhouette width", col=.silScale(x), show_column_names = FALSE,
        column_title = "Subpopulation mean silhouette width", row_names_max_width = unit(9,"cm"),
        bottom_annotation=.ds_anno(colnames(x)), heatmap_legend_param = list(direction="horizontal", legend_width=unit(3, "cm")))
p1 <- grid.grabExpr(draw(h1, heatmap_legend_side="bottom"))

res$clustering <- res$clustering[grep("norm\\.seurat\\.noscale.*denoise", row.names(res$clustering), invert=TRUE),]
h <- scrna_evalPlot_clust(res, what="MI", agg.by=c("norm","dr"), scale=TRUE, anno_legend = FALSE)
h <- h + scrna_evalPlot_clust(res, what="mean_F1", agg.by=c("norm","dr"), scale=TRUE, reorder_rows = h, anno_legend = FALSE) +
  scrna_evalPlot_clust(res, what="min_pr", agg.by=c("norm","dr"), scale=TRUE, reorder_rows = h, anno_legend = FALSE) +
  scrna_evalPlot_clustAtTrueK(res, agg.by=c("norm","dr"), reorder_rows = h, anno_legend = FALSE)
h <- .renameHrows(h, function(x) gsub("; dr=", " + ", gsub("seuratvst", "sctransform", x)))

p2 <- grid.grabExpr(draw(h))

plot_grid(p1,p2, nrow=2, labels=LETTERS[1:2])
```

**A:** Average silhouette width per subpopulation according to various normalization and dimension reductions. **B:** Clustering accuracy , mean proportion of the variance in the first 5 components explained by real subpopulations (center), and median ARI of the resulting clustering (right).