### 2020-01-30 supplementary_figures/supp6_clust.Rmd
added figure on nb of clusters, updated suppl

---
title: "Supplementary Figures relative to clustering"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r, include=FALSE}
if(!exists('FIG_NB')){
  FIG_NB <- 0
  getFigNb <- function(increment=FALSE){
    if(increment) FIG_NB <<- FIG_NB + 1
    FIG_NB
  }
}

suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(reshape2)
  library(dplyr)
  library(patchwork)
})
theme_set(theme_cowplot(font_size = 12))
devtools::load_all("../../pipeComp/")
source("../misc_functions.R")
```

# Supplementary Figure `r getFigNb(TRUE)`

```{r, include=FALSE}
#res <- readRDS("../data/clust_endSummary.rds")
res <- readRDS("../data/results_clust_2.rds")
res <- parsePipNames(res$clustering)
x <- grep("true\\.nbClusts",colnames(res),value=TRUE)
names(x) <- x <- sapply(strsplit(x," "),FUN=function(x) x[[1]])
aris <- as.data.frame(rbindlist(lapply(x, FUN=function(x){
  x <- res[,paste(x,"ARI"),drop=FALSE]
  colnames(x) <- "ARI"
  x
}), idcol = "dataset"))
aris$nbClusters <- unlist(lapply(x, FUN=function(x) res[,paste(x,"n_clus")]))
aris$clust.method <- res$clustmethod
aris$norm.method <- res$norm
levels(aris$norm.method) <- c("norm.seurat","sctransform")
aris <- aris[which(aris$nbClusters <= 40),]
dd <- data.frame( dataset=names(x), k=sapply(x, FUN=function(x) res[,paste(x,"true.nbClusts")][1]))
```

```{r, fig.width=7.5, fig.height=7}
ggplot(aris, aes(nbClusters, ARI, colour=clust.method, shape=norm.method)) + 
  geom_point() + xlim(c(1,40)) + facet_wrap(~dataset) + xlab("Number of clusters") +
  geom_vline(aes(xintercept=k), data=dd, linetype="dashed") + theme(legend.position="bottom") + 
  guides(colour=guide_legend(ncol=1, title = ""), shape=guide_legend(ncol=1, title=""))
```

### Supplementary Figure `r getFigNb()`
The number of clusters called has a much bigger impact on the Adjusted Rand Index (ARI) than differences between methods. The dashed line indicates the true number of clusters.

# Supplementary Figure `r getFigNb(TRUE)`

```{r, include=FALSE}
res <- res[which(res$norm=="norm.seurat" & res$resolution <5),]
x <- grep("true\\.nbClusts",colnames(res),value=TRUE)
names(x) <- x <- sapply(strsplit(x," "),FUN=function(x) x[[1]])
y <- cbind(res[,c("k","steps","resolution","clustmethod","dims")],sapply(x, FUN=function(x){
  res[[paste(x,"n_clus")]] - res[[paste(x,"true.nbClusts")]]
}))
y$param <- ifelse( res$clustmethod=="clust.seurat", paste0(" res=",res$resolution), paste0(" steps=",res$steps))
ag <- aggregate(y[,names(x)], by=y[,c("clustmethod", "k","param"),drop=FALSE], FUN=mean)
agl <- cbind(ag[rep(1:nrow(ag), 9),1:3], dataset=rep(colnames(ag)[4:ncol(ag)], each=nrow(ag)), delta=as.numeric(as.matrix(ag[,4:ncol(ag)])))
agl$k <- factor(agl$k, as.character(sort(unique(agl$k))))
x <- sapply(unique(strsplit(ag$param, "=")), FUN=function(x) as.numeric(x[[2]]) )
names(x) <- unique(ag$param)
agl$param <- factor(agl$param, names(sort(x)))

ssq <- trans_new("ssq", function(x) sign(x)*sqrt(abs(x)), function(x) sign(x)*x^2)
#slg <- trans_new("slg", function(x) sign(x)*log1p(abs(x)), function(x) sign(x)*(exp(x)-1))
pf <- function(x) ggplot(x, aes(k, param, fill=delta)) + geom_tile() + 
  facet_wrap(~dataset, nrow = 1) + labs(x="", fill="Difference to real number of clusters") + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank(), axis.line.x = element_blank()) +
  scale_fill_gradient2(low="darkblue", mid="white", high="red", na.value = "red", trans=ssq, limit=c(-7,20), breaks=c(-7,-2,0,2,7,20)) + 
  theme(legend.position = "none")
```

```{r, fig.width=8, fig.height=8}
(pf(agl[which(agl$clustmethod=="clust.seurat"),]) + ylab("Seurat") + theme(legend.position = "top")) / 
    (pf(agl[which(agl$clustmethod=="clust.scran"),]) + ylab("scran walktrap")) /
    (pf(agl[which(agl$clustmethod=="clust.scran.fg"),]) + ylab("scran fast_greedy")) /
    (ggplot(agl, aes(k)) + theme(axis.line.y=element_blank(), plot.margin = unit(rep(0,4),"pt"))) + 
    plot_layout(heights = c(10,8,8,1))
```

### Supplementary Figure `r getFigNb()`
Difference between the number of detected clusters and the number of real subpopulations according to different clustering paramters.

\newpage