### 2020-02-02 simulations/prepare_citeseq.Rmd
added code for simulations

---
title: "Prepare citeseq data for simulation"
author: "Pierre-Luc Germain"
date: "10/27/2019"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(SingleCellExperiment)
  library(scater)
  library(muscat)
})
source("pipeComp/inst/extdata/scrna_alternatives.R")
```

```{r}
download.file("ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE100nnn/GSE100866/suppl/GSE100866_CBMC_8K_13AB_10X-RNA_umi.csv.gz","RNA.csv.gz")
cbmc.rna <- as.sparse(read.csv(file = "RNA.csv.gz", sep = ",", header = TRUE, row.names = 1))
cbmc.rna <- cbmc.rna[-grep("ERCC_|MOUSE_",row.names(cbmc.rna)),]
sce <- SingleCellExperiment(list(counts=cbmc.rna))
sce <- addQCPerCell(sce)
out <- isOutlier(sce$detected, log=TRUE) | isOutlier(sce$sum, log=TRUE) | isOutlier(sce$percent_top_50)
sce <- sce[,!out]
cbmc.rna <- cbmc.rna[,colnames(sce)]

download.file("ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE100nnn/GSE100866/suppl/GSE100866_CBMC_8K_13AB_10X-ADT_umi.csv.gz", "ADT.csv.gz")
cbmc.adt <- as.sparse(read.csv(file = "ADT.csv.gz", sep = ",", 
    header = TRUE, row.names = 1))
cbmc.adt <- cbmc.adt[,colnames(sce)]
```

Then we more or less follow the Seurat procedure:

```{r}
cbmc <- CreateSeuratObject(counts = cbmc.rna)
cbmc <- NormalizeData(cbmc)
cbmc <- FindVariableFeatures(cbmc, nfeatures=2000)
cbmc <- ScaleData(cbmc)
cbmc <- RunPCA(cbmc, verbose = FALSE)
ElbowPlot(cbmc, ndims = 50)
cbmc <- FindNeighbors(cbmc, dims = 1:9)
cbmc <- FindClusters(cbmc, resolution = 0.8)
cbmc[["rnaClusterID"]] <- Idents(cbmc)
```

```{r}
cbmc[["ADT"]] <- CreateAssayObject(counts = cbmc.adt)
cbmc <- NormalizeData(cbmc, assay = "ADT", normalization.method = "CLR")
cbmc <- ScaleData(cbmc, assay = "ADT")
DefaultAssay(cbmc) <- "ADT"
adt.dist <- dist(t(GetAssayData(cbmc, slot = "data")))
cbmc[["adt_snn"]] <- FindNeighbors(adt.dist)$snn
cbmc <- FindClusters(cbmc, resolution = 0.2, graph.name = "adt_snn")
(tt <- table(Idents(cbmc), cbmc$rnaClusterID))
```

We then select cells of the table (i.e. cells clustered together on both types of data) for which there are enough cells:
```{r}
t2 <- t(apply(tt, 1, FUN=function(x) c(which.max(x)-1, max(x))))
t2 <- t2[order(t2[,2],decreasing=TRUE),]
t2 <- t2[!duplicated(t2[,1]) & t2[,2] > 200,]
to_keep <- paste(row.names(t2),t2[,1])
sce$cluster_id <- factor(paste(Idents(cbmc),cbmc$rnaClusterID))
sce2 <- sce[,sce$cluster_id %in% to_keep]
sce2$cluster_id <- droplevels(sce2$cluster_id)
levels(sce2$cluster_id) <- LETTERS[seq_along(levels(sce2$cluster_id))]
```

We prepare for simulation:
```{r}
sce2$sample_id <- factor(sample(LETTERS[1:2],ncol(sce2),replace=TRUE))
sce2 <- prepSim(sce2)
saveRDS(sce2, file="citeseq_prep4sim.SCE.rds")
```