### 2020-02-02 simulations/simulation.Rmd

---
title: "simulation"
author: "Pierre-Luc Germain"
date: "9/30/2019"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(muscat)
  library(purrr)
  library(scater)
  library(cowplot)
})
source("../../pipComp/R/misc.R")
source("../../pipComp/R/filtering.R")
source("../../pipComp/R/selection.R")
theme_set(theme_cowplot())
```

# Simulating dataset 1

```{r}
# see citeseq.Rmd
sce <- readRDS("citeseq_prep4sim.SCE.rds")
sce$sample_id <- factor("A")
```

```{r}
sims <- list()
set.seed(1234)
probs <- list( set_names(c(0.8,0.1,0.1,0,0),levels(sce$cluster_id)), 1, c(1,0))
lfcs <- set_names(c(2,3,2,1,1),levels(sce$cluster_id))
sims$s1 <- simData(sce, nrow(sce), 1000, probs=probs, p_type=0.2, rel_lfc=lfcs)
probs <- list( set_names(c(0.1,0.6,0,0.3,0), levels(sce$cluster_id)), 1, c(1,0))
lfcs <- set_names(c(2,1,1,1,1), levels(sce$cluster_id))
sims$s2 <- simData(sce, nrow(sce), 1000, probs=probs, p_type=0.05, rel_lfc=lfcs)
probs <- list( set_names(c(0.3,0.3,0.1,0,0.3), levels(sce$cluster_id)), 1, c(1,0))
lfcs <- set_names(rep(1,5), levels(sce$cluster_id))
sims$s3 <- simData(sce, nrow(sce), 500, probs=probs, p_type=0.01, rel_lfc=lfcs)

sim <- SingleCellExperiment(list(counts=do.call(cbind, lapply(sims,counts))))
sim$phenoid <- paste0( rep(LETTERS[1:3],sapply(sims,ncol)),
                       unlist(lapply(sims,FUN=function(x){
  levels(x$cluster_id) <- as.character(seq_along(levels(x$cluster_id)))
  x$cluster_id
})))
colnames(sim) <- paste0(sim$phenoid, ".c", 1:ncol(sim))
sim <- add_meta(sim)
sim <- compute_all_gene_info(sim)
saveRDS(sim, file="../datasets/simMix1.SCE.rds")
```

# Simulating dataset 2

```{r, eval=FALSE}
sce <- readRDS("~/bioinfo/robinson/muscat-comparison/MAGL/output/MAGL-SCE.rds")
sce$oldclusters <- sce$cluster_id
sce$cluster_id <- as.factor(gsub("[0-9]$","",sce$cluster_id))
tt <- table(sce$sample_id)
sce <- sce[,which(sce$sample_id %in% names(tt)[which(tt>3000)])]
sce$sample_id <- droplevels(sce$sample_id)
sce <- prepSim(sce,min_size=50)
saveRDS(sce, file="magl_sce_prepSim.rds")
```
```{r, include=FALSE}
sce <- readRDS("magl_sce_prepSim.rds")
```

```{r}
sce$offset <- sce$offset+log(2)
```

```{r}
sims <- list()
set.seed(1234)
probs1 <- list( set_names(c(0.2,0.4,0.35,0.05),levels(sce$cluster_id)),
               c(1,0), c(1,0) )
lfcs1 <- set_names(c(2,4,1,2),levels(sce$cluster_id))
sims$s1 <- simData(sce, nrow(sce), 1500, probs=probs1, p_type=0.1, rel_lfc=lfcs1)
probs2 <- list( set_names(c(0.2,0.6,0.2,0),levels(sce$cluster_id)),
               c(1,0), c(1,0) )
lfcs2 <- set_names(c(3,2,1,1),levels(sce$cluster_id))
sims$s2 <- simData(sce, nrow(sce), 1000, probs=probs2, p_type=0.03, rel_lfc=lfcs2)
probs3 <- list( set_names(c(0.5,0.5,0,0),levels(sce$cluster_id)),
               c(1,0), c(1,0) )
lfcs3 <- set_names(c(1,1,1,1),levels(sce$cluster_id))
sims$s3 <- simData(sce, nrow(sce), 500, probs=probs3, p_type=0.01, rel_lfc=lfcs3)

sim <- SingleCellExperiment(list(counts=do.call(cbind, lapply(sims,counts))))
sim$phenoid <- paste0( rep(LETTERS[1:3],c(1500,1000,500)),
                       unlist(lapply(sims,FUN=function(x){
  levels(x$cluster_id) <- as.character(seq_along(levels(x$cluster_id)))
  x$cluster_id
})))
colnames(sim) <- paste0(sim$phenoid, ".c", 1:ncol(sim))
sim <- add_meta(sim)
sim <- compute_all_gene_info(sim)
saveRDS(sim, file="../datasets/simMix2.SCE.rds")
```

# Overview

```{r, fig.width=8.5, fig.height=3}
ll <- c( simMix1="../datasets/simMix1.SCE.rds",
         simMix2="../datasets/simMix2.SCE.rds" )
devtools::load_all("../../pipeComp")
describeDatasets(ll)
```