### 2020-02-03 main_figures/normalization.Rmd

---
title: "Figure 3 normalization"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: true
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
  library(pipeComp)
})
source("../misc_functions.R")
theme_set(theme_cowplot(font_size = 11))
renameNorm <- function(x){
  .renameHrows(x, f=function(x){
    x <- gsub("^norm\\.", "", x)
    x <- gsub("seuratvst","sctransform",x)
    x <- gsub(".noscale", " (no scaling)", x)
  })
}
```

```{r}
res <- readRDS("../data/results_norm.rds")
```

```{r}
p1 <- scrna_evalPlot_DR(res, "covarRes", agg.by = "norm", anno_legend=FALSE, reorder_rows = )
p1 <- p1 + scrna_evalPlot_DR(res, "covarRes", covar = "log10_total_features", agg.by = "norm", anno_legend=FALSE, reorder_rows=p1)
p1 <- grid.grabExpr(draw(renameNorm(p1)))

p2 <- grid.grabExpr(draw(
  renameNorm(scrna_evalPlot_DR(res, what="silhouette", agg.by = "norm", value_format = "", show_heatmap_legend = TRUE))))

H <- scrna_evalPlot_clustAtTrueK(res, "ARI", agg.by = "norm", scale = TRUE, reorder_rows = TRUE, anno_legend=FALSE)
H <- H + scrna_evalPlot_clust(res, "MI", agg.by = "norm", scale = TRUE, reorder_rows=row.names(H@matrix), anno_legend=FALSE)
p3 <- grid.grabExpr(draw(renameNorm(H)))
```

```{r figure3, fig.width=9, fig.height=9}
plot_grid(p1,p2,p3, nrow = 3, labels=LETTERS[1:3])
```

Evaluation of normalization procedures. **A:** Residual correlation of the first principal component with library size (left) and detection rate (right), after accounting for biological differences between subpopulations. **B:** Average silhouette width per true subpopulation, where higher silhouette width means a higher separability. **C:** Clustering accuracy, measured by the ARI at the true number of cluster (left) and by the average mutual information (MI) of the cluster assignment with true subpopulations.