### 2020-02-27 supplementary_figures/supp6_clust.Rmd

---
title: "Supplementary Figures relative to clustering"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r, include=FALSE}
if(!exists('FIG_NB')){
  FIG_NB <- 0
  getFigNb <- function(increment=FALSE){
    if(increment) FIG_NB <<- FIG_NB + 1
    FIG_NB
  }
}

suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(reshape2)
  library(dplyr)
  library(patchwork)
  library(pipeComp) # (requires `pipeComp` version >=0.99.3)
  library(scales)
})
theme_set(theme_cowplot(font_size = 12))
source("../misc_functions.R")
```

# Supplementary Figure `r getFigNb(TRUE)`

```{r, include=FALSE}
res <- readRDS("../data/results_clustering2.rds")
res <- res$evaluation$clustering
res <- res[which(res$norm=="norm.seurat" & res$resolution <5),]
res$delta <- res$n_clus-res$true.nbClusts
res$param <- ifelse( res$clustmethod=="clust.seurat", paste0(" res=",res$resolution), paste0(" steps=",res$steps))
res$param[which(res$clustmethod=="clust.scran.fg")] <- "NA"
agl <- aggregate(res[,"delta",drop=FALSE], by=res[,c("dataset","clustmethod", "k","param"),drop=FALSE], FUN=mean)
agl$k <- factor(agl$k, as.character(sort(unique(agl$k))))
x <- sapply(unique(strsplit(agl$param, "=")), FUN=function(x){
  tryCatch(as.numeric(x[[2]]), error=function(e) NA_real_)
})
names(x) <- unique(agl$param)
agl$param <- factor(agl$param, names(sort(x)))

ssq <- trans_new("ssq", function(x) sign(x)*sqrt(abs(x)), function(x) sign(x)*x^2)
#slg <- trans_new("slg", function(x) sign(x)*log1p(abs(x)), function(x) sign(x)*(exp(x)-1))
pf <- function(x) ggplot(x, aes(k, param, fill=delta)) + geom_tile() + 
  facet_wrap(~dataset, nrow = 1) + labs(x="", fill="Difference to real number of clusters") + 
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank(), axis.line.x = element_blank()) +
  scale_fill_gradient2(low="darkblue", mid="white", high="red", na.value = "red", trans=ssq, limit=c(-7,20), breaks=c(-7,-2,0,2,7,20)) + 
  theme(legend.position = "none")
```

```{r, fig.width=8, fig.height=8}
tt <- theme(strip.background = element_blank(), strip.text.x = element_blank())
p1 <- pf(agl[which(agl$clustmethod=="clust.seurat"),]) + ylab("Seurat") + 
  theme(legend.position = "top", strip.text.x = element_text(size = 10, angle=60), strip.background = element_blank())
p1 / 
    (pf(agl[which(agl$clustmethod=="clust.scran"),]) + ylab("scran walktrap") + tt) /
    (pf(agl[which(agl$clustmethod=="clust.scran.fg"),]) + ylab("scran fast_greedy") + tt) /
    (ggplot(agl, aes(k)) + theme(axis.line.y=element_blank(), plot.margin = unit(rep(0,4),"pt"))) + 
    plot_layout(heights = c(10,6,2,1))
```

### Supplementary Figure `r getFigNb()`
Difference between the number of detected clusters and the number of real subpopulations according to different clustering paramters.

\newpage