### 2020-03-01 main_figures/normalization.Rmd
updated normalization figure

---
title: "Normalization"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
  library(pipeComp)
})
source("../misc_functions.R")
theme_set(theme_cowplot(font_size = 10))
```

(Requires `pipeComp` version >=0.99.3)

```{r}
res <- readRDS("../data/results_norm.rds")
```

```{r}
p1 <- scrna_evalPlot_DR(res, "log10_total_counts", agg.by = "norm", anno_legend=FALSE)
p1 <- p1 + scrna_evalPlot_DR(res, "log10_total_features", agg.by = "norm", anno_legend=FALSE, reorder_rows=p1)
p1 <- grid.grabExpr(draw(renameNorm(p1)))

p2 <- grid.grabExpr(draw(
  renameNorm(scrna_evalPlot_DR(res, what="meanSilWidth", agg.by = "norm", value_format = "", show_heatmap_legend = TRUE))))

H <- scrna_evalPlot_clust(res, "MI", agg.by = "norm", scale = TRUE, reorder_rows=TRUE, anno_legend=FALSE)
H <- H + scrna_evalPlot_clust(res, "ARI", agg.by = "norm", scale = TRUE, reorder_rows = row.names(H@matrix), anno_legend=FALSE) +
  scrna_evalPlot_clust(res, "ARI", atTrueK = TRUE, agg.by = "norm", scale = TRUE, reorder_rows = row.names(H@matrix), anno_legend=FALSE)
p3 <- grid.grabExpr(draw(renameNorm(H)))
```

```{r figure3, fig.width=9, fig.height=11}
pdf("normalization.pdf", width=10, height=11)
plot_grid(p1,p2,p3, nrow = 3, labels=LETTERS[1:3], rel_heights = c(10,9,10))
dev.off()
plot_grid(p1,p2,p3, nrow = 3, labels=LETTERS[1:3], rel_heights = c(10,9,10))
```

Evaluation of normalization procedures. **A:** Average of the absolute within-subpopulation correlation of the first 5 principal components with library size (left) and detection rate (right). **B:** Average silhouette width per true subpopulation, where higher silhouette width means a higher separability. **C:** Clustering accuracy, measured by the mutual information (MI), the adjusted Rand Index (ARI), and the ARI at the true number of cluster.