### 2020-04-20 svadea/run.Rmd

---
title: "SVA-DEA pipeline run"
author: "Pierre-Luc Germain"
date: "4/14/2020"
output: html_document
---

## Running the pipeline

```{r}
suppressPackageStartupMessages({
  library(SummarizedExperiment)
#  library(pipeComp)
})
devtools::load_all("../pipeComp/")
#source(system.file("extdata", "dea_wrappers.R", package="pipeComp"))
source("../pipeComp/inst/extdata/dea_wrappers.R")

pd <- dea_pipeline()

ds <- list.files("data_preparation", pattern="\\.rds$", full.names = TRUE)
names(ds) <- gsub("\\.rds$","",basename(ds))

none <- function(x, ...)  return(x)

alternatives <- list(
  filt=c("none","def.filter"),
  minCount=c(10,20),
  sva.method=c("none","sva.svaseq","sva.vstsva","sva.RUVr","sva.RUVs"),
  k=c(1:5,"be","leek"),
  dea.method=c("dea.edgeR", "dea.edgeR.QLF", "dea.DESeq2", "dea.voom")
)
eg <- buildCombMatrix(alternatives)
eg <- eg[eg$sva.method %in% c("sva.svaseq", "sva.vstsva") | 
           !(eg$k %in% c("be","leek")),]
eg <- eg[eg$filt=="def.filter" | eg$minCount==10,]
eg <- eg[eg$k=="1" | eg$sva.method!="none",]

res <- runPipeline(ds[2:3], alternatives, pd, output.prefix = "pipelineResults_test/", debug=FALSE, nthreads = 4)
```

## Results

```{r}
dea_evalPlot_logFCs(res, what=c("logFC.pearson","logFC.spearman"), agg.by=c("sva.method","dea.method"), reorder_rows=FALSE)
dea_evalPlot_curve(res, agg.by=c("sva.method","dea.method"), colourBy="sva.method", shapeBy="dea.method")
dea_evalPlot_sig(res, agg.by=c("sva.method","dea.method"), reorder_rows=FALSE)
dea_evalPlot_curve(res, agg.by=c("sva.method")) + ggtitle("Averaging across DEA methods")
dea_evalPlot_curve(res, agg.by=c("dea.method")) + ggtitle("Averaging across SVA strategies")
```

## Session info

```{r}
sessionInfo()
```