### 2020-05-04 main_figures/clustering.Rmd
updated many figures to improve readability...

---
title: "Clustering"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(pipeComp) # version >= 0.99.26
})
source("../misc_functions.R")
theme_set(theme_cowplot(font_size = 12))
```

```{r clustering_accuracy, fig.width=12, fig.height=3}
# biorxiv data & figure
# res <- readRDS("../data/results_clust_2.rds")
# h <- scrna_evalPlot_clust(res, what="MI", agg.by=c("norm","clustmethod"), scale=TRUE, anno_legend = FALSE)
# h <- h + scrna_evalPlot_clust(res, what="mean_F1", agg.by=c("norm","clustmethod"), scale=TRUE, reorder_rows = h, anno_legend = FALSE) +
#   scrna_evalPlot_clust(res, what="min_pr", agg.by=c("norm","clustmethod"), scale=TRUE, reorder_rows = h, anno_legend = FALSE) +
#   scrna_evalPlot_clustAtTrueK(res, agg.by=c("norm","clustmethod"), reorder_rows = h, anno_legend = FALSE)
# h <- pipeComp:::.renameHrows(h, function(x) gsub("seurat; clustmethod=|; clustmethod=", "", gsub("seuratvst", "sctransform + ", x)))
# h <- pipeComp:::.renameHrows(h, function(x) gsub("clust.scran","scran (walktrap)",gsub("clust.scran.fg", "scran (fast_greedy)", x)))

res <- readRDS("../data/results_clustering.rds")
cl <- res$evaluation$clustering
cl$graph <- cl$clustmethod
levels(cl$graph) <- c(paste0("scran (",c("exact rank SNN", "Annoy rank SNN", "exact rank SNN", "exact KNN", "Annoy KNN", "exact SNN", "Annoy SNN"),")"), "Seurat SNN")
levels(cl$clustmethod) <- c("walktrap","walktrap","fast-greedy","walktrap","walktrap","walktrap","walktrap","Seurat")
cl$elapsed <- res$elapsed$stepwise$clustering$elapsed
cl$nbClust.diff <- cl$n_clus-cl$true.nbClusts
clo <- cl
cl <- cl[cl$k<=20,]
res$evaluation$clustering <- cl

renameFn <- function(x){
  x <- gsub("seurat; graph=|; graph=", "", gsub("seuratvst", "sctrans.+", x))
  x <- gsub("; clustmethod=", "+", x)
  x <- gsub("^graph=","",x)
  x <- gsub("exact ","",x)
  x <- gsub("+walktrap","",x,fixed=TRUE)
}

tmp <- scale(pipeComp:::.prepRes(cl, "ARI", agg.by=c("norm","graph","clustmethod"))) + 
  scale(pipeComp:::.prepRes(cl, "MI", agg.by=c("norm","graph","clustmethod")))
o <- names(sort(rowMeans(tmp, na.rm=TRUE)+apply(tmp,1,na.rm=TRUE,FUN=median), decreasing = TRUE))
o <- gsub("norm=norm\\.","",o)
o <- c(o[grep("seuratvst",o)],o[grep("seuratvst",o,invert=TRUE)])

h <- renameHrows(
  evalHeatmap(res, step="clustering", what=c("MI","ARI","min_pr"), row_split = norm, font_factor = 0.85,
              agg.by=c("norm","graph","clustmethod"), anno_legend=FALSE) +
  evalHeatmap(res, step="clustering", what="ARI", agg.by=c("norm","graph","clustmethod"), font_factor = 0.85, filter=n_clus==true.nbClusts, title="ARI at\ntrue k", name="ARIatk",anno_legend=FALSE), function(x) sapply(strsplit(x," > "), FUN=function(x) paste(x[-1],collapse=" > ")))

h2 <- grid.grabExpr(draw(h))
```

```{r clustering_time}
clo$elapsed <- res$elapsed$stepwise$clustering$elapsed
# keep reasonable parameters for a realistic time
res$evaluation$clustering <- clo[which(clo$k<=20),]
h3 <- evalHeatmap(res, step="clustering", what="elapsed", agg.by=c("graph","clustmethod"), font_factor = 0.85,
                           anno_legend = TRUE, scale=TRUE, value_format = "%.1f", reorder_rows=TRUE,
                  row_names_max_width = unit(8, "cm"))
h3 <- h3 + evalHeatmap(res, step="clustering", what="nbClust.diff", 
                       row_names_max_width = unit(8, "cm"),
                        agg.by=c("graph","clustmethod"), agg.fn = median, font_factor = 0.85,
                        reorder_rows=h3, scale = FALSE, value_format = "%.1f", 
                        col = circlize::colorRamp2(c(-5, 0, 5), c("red", "white", "Darkblue")),
                        value_cols = c("black","black") )
h3 <- grid.grabExpr(draw(h3))
```

```{r clustering, fig.width=11, fig.height=8}
pdf("clustering.pdf", width=11.5, height=8)
plot_grid(h2, h3, labels=LETTERS[1:2], ncol=1, rel_heights = c(7,4))
dev.off()
plot_grid(h2, h3, labels=LETTERS[1:2], ncol=1, rel_heights = c(7,4))
```
Clustering accuracy (A), computing time (B left) and mean deviation from the true number of clusters (B right). The (exact) rank-based SNN graph represents the default setting in `scran`-based clustering; however we find that the Annoy approximation to the KNN diminishes the computing costs without negatively affecting the accuracy.