### 2020-05-04 main_figures/normalization.Rmd

---
title: "Normalization"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
  library(pipeComp)
})
source("../misc_functions.R")
theme_set(theme_cowplot(font_size = 10))
```

(Requires `pipeComp` version >=0.99.26)

```{r}
res <- readRDS("../data/results_norm.rds")
```

```{r}
classifyNorms <- function(x){
  y <- rep("other", length(x))
  y[grep("sctransform",x)] <- "sctransform"
  y[grep("seurat",x)] <- "seurat"
  y[grep("scran",x)] <- "scran"
  y[grep("stableG",x)] <- "stableG"
  y
}

p1 <- evalHeatmap(res, step="dimreduction", what=c("log10_total_counts","log10_total_features"), 
                  what2="PC1.covar.adjR2", agg.by = "norm", reorder_rows = TRUE, 
                  row_split = classifyNorms(norm), row_title="Variance in PC1 explained by covariate")
lgd <- Legend(col_fun=colorRamp2(1:100, viridisLite::inferno(100)), at=c(1,50,100), 
              title="MADs", labels=c("lowest","median","highest"))
p1 <- grid.grabExpr(draw(renameNorm(p1), merge_legend=TRUE, annotation_legend_list=lgd))

p2 <- grid.grabExpr(draw(renameNorm(
  scrna_evalPlot_silh(res, reorder_rows = TRUE, row_split = classifyNorms(norm),
                      row_title="Subpopulations silhouette width", anno_legend=FALSE)
  ), annotation_legend_side="right"))

lgd <- Legend(col_fun=colorRamp2(1:100, viridisLite::inferno(100)), at=c(1,50,100), 
              title="MADs", labels=c("worst","median","best"), direction="horizontal")
p3 <- grid.grabExpr(draw(
  renameNorm(evalHeatmap(res, step="clustering", what=c("MI","ARI"), agg.by = "norm", 
              reorder_rows=TRUE, anno_legend=FALSE, row_split=classifyNorms(norm), row_title="Clustering accuracy") + 
    evalHeatmap(res, step="clustering", what="ARI", agg.by = "norm", 
                filter=n_clus==true.nbClusts, reorder_rows=TRUE, anno_legend=FALSE,
                row_split=classifyNorms(norm), name="ARI at true k", title="ARI at true k")),
  heatmap_legend_side = "bottom", merge_legend=TRUE, annotation_legend_list=lgd))
```

```{r figure3, fig.width=9, fig.height=11}
pdf("normalization.pdf", width=10, height=11)
  plot_grid(p1,p2,p3, nrow = 3, labels=LETTERS[1:3], rel_heights = c(9,9,10))
dev.off()
plot_grid(p1,p2,p3, nrow = 3, labels=LETTERS[1:3], rel_heights = c(9,9,10))
```

Evaluation of normalization procedures. **A:** Variance in the first principal component explained by 
library size and detection rate after accounting for subpopulation differences. The values are difference in adjusted _R^2_ between a `PC1~subpopulation+covariate` model and `PC1~subpopulation`. **B:** Minimum and mean silhouette width per true subpopulation, where higher silhouette width means a higher separability. **C:** Clustering accuracy, measured by the mutual information (MI), the adjusted Rand Index (ARI), and the ARI at the true number of cluster. The coloring scheme in panels **A** and **B** is based on the per-metric number of median absolute deviations (MADs) from the per-column medians, and mapped to color using a square-root transformation to minimize the effect of outliers. The values shown in the cells are the direct metric values (no transformation except rounding).