### 2020-05-04 main_figures/imputation.Rmd
updated imputation figure with new color mapping

---
title: "Imputation/denoising"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
  library(pipeComp)
})
source("../misc_functions.R")
theme_set(theme_cowplot(font_size = 10))
```

```{r}
res <- readRDS("../data/results_imputation.rds")
dr <- res$dr
dr <- dr[order(rowMeans(dr>0,na.rm=TRUE)+rowMedians(dr, na.rm=TRUE), decreasing=TRUE),order(colMeans(dr,na.rm=TRUE),decreasing=TRUE)]
ds <- factor(sapply(strsplit(colnames(dr),"\\."),FUN=function(x) x[[1]]))
cols <- getQualitativePalette(length(unique(ds)))
names(cols) <- sort(unique(ds))
dsa <- ComplexHeatmap::HeatmapAnnotation(dataset=ds, col=list(dataset=cols), show_annotation_name = FALSE, show_legend=TRUE)

h1 <- Heatmap(dr, cluster_columns = FALSE, col=pipeComp:::.silScale(dr), cluster_rows = FALSE, 
        bottom_annotation=dsa,
        show_column_names = FALSE, show_column_dend = FALSE, 
        name="silh.\nwidth", column_title="Mean silhouette width", 
        right_annotation=rowAnnotation( mean=rowMeans(dr, na.rm=TRUE), 
                                        median=rowMedians(dr, na.rm=TRUE), 
                                        min=rowMins(dr, na.rm=TRUE), 
                                        show_legend=FALSE)
        )
p1 <- grid.grabExpr(draw(h1))

cl <- res$cl
mycast <- function(x,y){
  x <- reshape2::dcast(x, method~dataset, value.var=y, fun.aggregate = mean, na.rm=TRUE)
  row.names(x) <- x[,1]
  as.matrix(x[,-1])
}
ag <- aggregate(cl[,c("ARI",grep("_",colnames(cl),value=TRUE))], by=cl[,1:2], FUN=function(x){
  mean(as.numeric(unlist(x)),na.rm=TRUE)
  })
clw <- cl[which(cl$nbClustDeviation==0),]
ag2 <- aggregate(clw[,c("ARI",grep("_",colnames(clw),value=TRUE))], by=clw[,1:2], FUN=function(x){
  mean(as.numeric(unlist(x)),na.rm=TRUE)
})

ll2 <- list( "mean ARI"=mycast(ag, "ARI"),
            "ARI at true\nnb clusters"=mycast(ag2, "ARI"),
            "Mean precision"=mycast(ag, "mean_pr"),
            "Mean recall"=mycast(ag, "mean_re")
            )

e <- do.call(cbind, ll2)
o <- names(sort(rowMeans(e, na.rm=TRUE)+rowMedians(e, na.rm=TRUE), decreasing=TRUE))
ll2 <- lapply(ll2, FUN=function(x) x[o,])
ll2s <- lapply(ll2, pipeComp::colCenterScale)

myh <- function(k,title=k,...){
  rmLead0 <- sum(abs(ll2[[k]])>1,na.rm=TRUE)==0
  sp <- paste0('sprintf("%.2f", ll2[["',k,'"]][i, j])')
  resmid <- range(ll2s[[k]], na.rm=TRUE)
  resmid <- resmid[1]+(resmid[2]-resmid[1])/2
  cellfn <- eval(parse(text=paste0('function(j, i, x, y, width, height, fill){',
      'lab <- ', ifelse(rmLead0, paste('gsub("^0\\\\.",".",', sp,')'), sp),
      '; grid.text(lab, x, y, gp = gpar(fontsize = 10, 
      col=ifelse(ll2s[["',k,'"]][i,j]>',resmid,',"black","white"))) }')))
  Heatmap(ll2s[[k]], col = viridis::inferno(100), show_column_names = FALSE, bottom_annotation = pipeComp:::.ds_anno(colnames(ll2[[k]]), FALSE), cluster_columns=FALSE, cluster_rows = FALSE, show_heatmap_legend = FALSE, column_title=title, cell_fun=cellfn, ...)
}

p2 <- grid.grabExpr(draw( myh(names(ll2)[1]) + myh(names(ll2)[2]) + 
                            myh(names(ll2)[3]) + myh(names(ll2)[4]) ))
```

```{r imputation, fig.width=10, fig.height=7}
pdf("imputation.pdf", width = 10, height=7)
plot_grid( p1, p2, nrow=2, labels=LETTERS[1:2], rel_heights = c(4.5,6))
dev.off()

plot_grid( p1, p2, nrow=2, labels=LETTERS[1:2], rel_heights = c(4,6))
```

**A:** Average silhouette width per subpopulation across various denoising/imputation methods. **B:** Clustering accuracy across denoising/imputation methods.