### 2020-05-08 main_figures/dimensionality.Rmd
some figure updates

---
title: "Dimensionality"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
  library(pipeComp) # the first part requires pipeComp >0.99.26
})
source("../misc_functions.R")
theme_set(theme_cowplot(font_size = 12))
```

```{r, fig.width=8, fig.height=4}
rnfn <- function(x){
  x <- gsub("norm\\.","",x)
  x <- gsub("maxLikGlobal$","maxLikGlobal(k=20)",x)
  gsub("maxLikGlobal10$","maxLikGlobal(k=10)",x)
}
res <- readRDS("../data/results_dimensionality.rds")
H <- renameHrows(
  scrna_evalPlot_silh( res, step="dimensionality", reorder_rows=TRUE, 
                       row_split=gsub("norm\\.","",norm),
                       heatmap_legend_param=list(direction="horizontal")),
  rnfn)
p2 <- grid.grabExpr(draw(H, heatmap_legend_side="bottom"))

p3 <- renameHrows(
  evalHeatmap( res, what=c("NMI","min_pr"), agg.by=c("norm","dims"), 
                   row_split=gsub("norm\\.","",norm), reorder_rows=TRUE, anno_legend=FALSE) + 
  evalHeatmap(res, what=c("ARI"), agg.by=c("norm","dims"), row_split=gsub("norm\\.","",norm), 
              reorder_rows=TRUE, filterExpr=n_clus==true.nbClusts, 
              anno_legend=FALSE, title="ARI at\ntrue k"),
  rnfn
)
p3 <- grid.grabExpr(draw(p3))
```

```{r, eval=FALSE}
# NB: the next results were generated using an older version of pipeComp (branch 
# `biorxiv`), and must be plotted using this version of the package
suppressPackageStartupMessages({
  library(pipeComp)
})
```

```{r, include=FALSE}
devtools::load_all("../../pipeComp_biorxiv/pipeComp/")
source("../../pipeComp_biorxiv/pipeComp/inst/extdata/scrna_alternatives.R")
```

```{r}
o <- readRDS("../data/ndims.RDS")
o <- lapply(o, FUN=function(x){
  x$nselected$elbow <- x$nselected$elbow-1
  x
})
o2 <- data.table::rbindlist( lapply(o, FUN=function(x){
  data.frame( method=names(x$nselected), 
              nsel=unlist(x$nselected), 
              time=x$elapsed )
  }), idcol="dataset")
ve <- sapply(o, FUN=function(x) x$varExplained)
row.names(ve) <- 1:nrow(ve)

cols <- getQualitativePalette(9)[-5:-6]
names(cols) <- colnames(ve)
p1a <- Heatmap(t(ve)[,1:15],col=c("white","black"), cluster_columns=FALSE, cluster_rows=FALSE,name="Prop.\nvar", column_title = "Variance in each PC\nexplained by subpopulations", row_title = "Dataset", right_annotation = rowAnnotation(dataset=colnames(ve), col=list(dataset=cols), show_legend=FALSE), show_row_names = FALSE)
p1a <- grid.grabExpr(draw(p1a))

ve.elbows <- apply(ve,2,FUN=farthestPoint)-1
o2$diff <- o2$nsel-ve.elbows[as.character(o2$dataset)]
tmp <- sapply(split(o2$time, o2$method), mean)
mt2 <- cut(tmp, breaks=c(0,1,5,20,3000), labels=c("<1s","<5s","<20s","minutes \nto hours"))
names(mt2) <- names(tmp)
o2$Time <- mt2[as.character(o2$method)]
tcols <- c("darkblue", "blue", "violet", "red")
names(tcols) <- levels(mt2)
o2$avgTime <- tmp[as.character(o2$method)]
p1b <- ggplot(o2, aes(reorder(method, diff, function(x) mean(abs(x))), diff, group=method, fill=Time)) + geom_violin() + coord_flip() + geom_hline(yintercept=0, linetype="dashed") + ylab("Difference to expected dimensionality") + scale_fill_manual(values=tcols) + theme(axis.title.y=element_blank(), legend.position=c(0.7, 0.3))
p1 <- plot_grid(p1a, p1b, rel_widths = c(0.42,0.58), labels=LETTERS[1:2])
```

```{r figure6, fig.width=9, fig.height=9.5}
pdf("dimensionality.pdf", width=9, height=10)
plot_grid( p1, p2, p3, nrow=3, labels=c(NA,LETTERS[3:4]))
dev.off()
plot_grid( p1, p2, p3, nrow=3, labels=c(NA,LETTERS[3:4]))
```

**A:** Estimated dimensionality using the proportion of variance in each component explained by subpopulations. **B:** Difference between `real' dimensionality (from **A**) and various dimensionality estimation methods, along with computing time. **C:** Average silhouette width per subpopulation across a selection of dimensionality estimators, using sctransform or standard seurat normalization. **D:** Clustering accuracy of across the normalization/dimensionality methods.