### 2020-05-08 data/overall_run.Rmd
added run info for the overall benchmark

---
title: "Overall run"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(SingleCellExperiment)
  library(Matrix)
  library(devtools)
  library(scater)
  library(pipeComp)
})
source(system.file("extdata", "scrna_alternatives.R", package="pipeComp"))
ds_files <- list.files("datasets", full.names = T)
names(ds_files) <- gsub(".SCE.rds","",basename(ds_files),fixed=T)
names(ds_files)
```

## Methods

```{r}
clust.scran.knnAnnoy <- function(ds, ...){
  clust.scran(ds, graph.type="knn", neighbor.method="Annoy", ...)
}
none <- function(x, ...) return(x)

pd <- scrna_pipeline(pipeClass="seurat")

alternatives <- list(
  doubletmethod=c("none","doublet.scDblFinder"),
  filt=c("none","filt.default","filt.stringent"),
  norm=c("norm.seurat","norm.sctransform","norm.scran"),
  sel=c("sel.vst","sel.deviance","sel.expr"),
  selnb=c(2000,4000),
  dr=c("seurat.pca"),
  maxdim=50,
  dims=c("elbow", 10, "maxLikGlobal"),
  clustmethod=c("clust.seurat","clust.scran.knnAnnoy"),
  k=c(5,8,10,15,20,30),
  steps=4,
  resolution=c(0.001,0.005,0.01,0.02,0.05,0.1,0.15,0.2,0.3,0.4,0.5,0.8,1,1.2,1.5,2,4),
  min.size=20
)

comb <- buildCombMatrix(alternatives)
comb <- comb[(comb$clustmethod=="clust.seurat" | comb$resolution==0.8),]
```

## Run

```{r}
res <- runPipeline(ds_files, alternatives, pd, output.prefix="pipeComp_overall/", nthreads=18, comb=comb)
saveRDS(res, file="pipeComp_overall/summary.rds")
```

```{r session_info}
sessionInfo()
```