### 2020-05-13 main_figures/svadea.Rmd

---
title: "SVA-DEA benchmark"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(pipeComp)
  library(grid)
  library(ComplexHeatmap)
  library(circlize)
})
source("../misc_functions.R")
theme_set(theme_cowplot(font_size = 11))
res1 <- res <- readRDS("../svadea/results/aggregated.rds")
res1$evaluation$dea <- lapply(res1$evaluation$dea, FUN=function(x){
  x[which(x$k==1),]
})
```

```{r, fig.width=11.5, fig.height=6.8}
# hts <- dea_evalPlot_sig(res, what=c("logFC.pearson","TPR"), agg.by=c("sva.method","k"), row_split="sva.method", 
#                      anno_legend=FALSE, show_column_names=TRUE, font_factor=0.85, scale="colCenterScale",
#                      show_heatmap_legend=TRUE, heatmap_legend_param=list(direction="horizontal") ) +
#   dea_evalPlot_sig(res, what="FDR", agg.by=c("sva.method","k"), row_split="sva.method", 
#                      anno_legend=FALSE, show_column_names=TRUE, font_factor=0.85, scale="colCenterScale",
#                      col=rev(viridisLite::inferno(100)), value_cols=c("white","black"), 
#                      show_heatmap_legend=TRUE, heatmap_legend_param=list(direction="horizontal") )
# draw(hts, heatmap_legend_side = "bottom", merge_legend=TRUE)

hts <- evalHeatmap( res, step="dea", what="logFC.pearson", agg.by=c("sva.method","k"), 
                    row_split="sva.method", anno_legend=FALSE, show_column_names=TRUE, 
                    font_factor=0.85 ) +
  evalHeatmap( res, step="dea", what="TPR", agg.by=c("sva.method","k"), 
                    row_split="sva.method", anno_legend=FALSE, show_column_names=TRUE, 
                    font_factor=0.85, filterExpr=threshold==0.05 ) +
  evalHeatmap(res, step="dea", what="FDR", agg.by=c("sva.method","k"), row_split="sva.method", 
                     anno_legend=FALSE, show_column_names=TRUE, font_factor=0.85, 
                     col=rev(viridisLite::inferno(101)), value_cols=c("white","black"), filterExpr=threshold==0.05)
hts <- .renameHrows(hts, f = function(x) gsub("none > 1", "none",x, fixed=TRUE))
#draw(hts, heatmap_legend_side = "bottom", merge_legend=TRUE, annotation_legend_list=.scaledLegend())

p <- plot_grid(
  plot_grid(
    dea_evalPlot_curve(res1, agg.by=c("sva.method","dea.method"), colourBy="dea.method", shapeBy="sva.method") + ggtitle("Using a single surrogate variable"),
    dea_evalPlot_curve(res, agg.by=c("sva.method","k"), colourBy="k", shapeBy="sva.method") + ggtitle("Increasing the number of surrogate variables"),
    nrow=2, labels=LETTERS[1:2]
  ),
  grid.grabExpr(draw( hts, heatmap_legend_side = "bottom", merge_legend=TRUE, 
                      annotation_legend_list=.scaledLegend())),
  nrow=1, rel_widths=c(5,2.2), labels=c(NA,"C")
)
pdf("svadea.pdf", width=11.5, height=6.8)
p
dev.off()
p
```