### 2020-06-18 main_figures/overall.Rmd
added results-overview figure

---
title: "Overall"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(ComplexHeatmap)
  library(S4Vectors)
  library(matrixStats)
  library(pipeComp) # version >= 0.99.26
})
source("../misc_functions.R")
```

```{r}
##########
## Note that a very similar plot can be created much more simply using the
## `scrna_evalPlot_overall` function of pipeComp. Here we do it manually in
## order to customize the positions of legends and such.

ov <- readRDS("../data/overall.rds")

ll <- lapply(c("minSilWidth", "meanSilWidth"), FUN=function(x){
  pipeComp:::.prepRes(ov$evaluation$dimreduction$silhouette$top_10_dims, what=x,
                      agg.by=c("doubletmethod","filt","norm","sel"), 
                      pipDef=metadata(ov)$PipelineDefinition, 
                      returnParams=TRUE, shortNames=TRUE)
})

pp1 <- ll[[1]]$pp
pp1$method <- NULL
ll1 <- lapply(ll, FUN=function(x) x$res)

ll <- lapply(c("ARI", "MI"), FUN=function(x){
  pipeComp:::.prepRes(ov$evaluation$clustering, what=x,
                      agg.by=c("doubletmethod","filt","norm","sel","dims","clustmethod"), 
                      pipDef=metadata(ov)$PipelineDefinition, 
                      returnParams=TRUE, shortNames=TRUE)
})
ll[[3]] <- pipeComp:::.prepRes(ov$evaluation$clustering, what="ARI",
                      agg.by=c("doubletmethod","filt","norm","sel","dims","clustmethod"), 
                      pipDef=metadata(ov)$PipelineDefinition, 
                      returnParams=TRUE, shortNames=TRUE, filt=rlang::expr(true.nbClusts==n_clus))
pp <- ll[[1]]$pp
pp$method <- NULL
ll2 <- lapply(ll, FUN=function(x) colCenterScale(x$res))

pclost <- scrna_evalPlot_filtering(ov, returnTable = TRUE)
pclost <- reshape2::dcast(pclost, doubletmethod+filt~dataset, value.var="max.lost")
row.names(pclost) <- apply(pclost[,1:2],1,collapse=" > ",FUN=paste)
tmp <- as.character(apply(pp[,colnames(pclost)[1:2]],1,collapse=" > ",FUN=paste))
pclost <- pclost[tmp,-1:-2]
row.names(pclost) <- row.names(ll2[[1]])
pclost <- apply(pclost,1,FUN=max)

tmp <- as.character(apply(pp[,colnames(pp1)],1,collapse=" > ",FUN=paste))
ll1 <- lapply(ll1, FUN=function(x){ x <- x[tmp,]; row.names(x) <- row.names(ll2[[1]]); x})
ll <- c(ll1,ll2)
```

```{r}
ll2 <- list( list(mat=ll[[1]], title="min silhouette\nwidth", cluster_columns=TRUE, name="silhouette width"),
             list(mat=ll[[2]], title="mean silhouette\nwidth", cluster_columns=TRUE, show_heatmap_legend=FALSE),
             list(mat=ll[[3]], title="mean ARI", name="ARI (MADs)", cluster_columns=FALSE),
             list(mat=ll[[4]], title="mean MI", name="MI (MADs)", cluster_columns=FALSE),
             list(mat=ll[[5]], title="mean ARI at\ntrue k", name="ARI at true k (MADs)", cluster_columns=FALSE) )
silhscale <- pipeComp:::.silScale(cbind(ll2[[1]]$mat, ll2[[2]]$mat))

ro <- rowMeans(do.call(cbind, lapply(ll[c(2:4)], FUN=function(x){
  x/sd(x,na.rm=TRUE)
  rowMeans(x,na.rm=TRUE)+rowMins(x,na.rm=TRUE)
})))
names(ro) <- row.names(ll[[1]])
h <- hclust(dist(do.call(cbind,ll)))
h2 <- reorder(as.dendrogram(h), ro)
ro2 <- names(sort(ro,decreasing=TRUE))

pp$doubletmethod <- gsub("doublet\\.","",pp$doubletmethod)
pp$filt <- factor(gsub("filt\\.","",pp$filt), c("none","default","stringent"))
pp$sel <- gsub("sel\\.","",pp$sel)
pp$norm <- gsub("norm\\.","",pp$norm)
pp$clustmethod <- gsub("clust\\.","",pp$clustmethod)
colnames(pp) <- c("doubet","filtering","normalization","feature sel","# dimensions","clustering")

cols <- getQualitativePalette(15)
vals <- as.character(unlist(lapply(lapply(pp, FUN=unique), as.character)))
cols <- c("none"="white", "scDblFinder"="black", "stringent"="black", "default"="darkgrey", sctransform="#114477", scran="#117755", seurat="#AA4444", deviance="#777711", vst="#771111", expr="#DD77AA", "maxLikGlobal"="#77AADD", "10"="#AAAA44", elbow="#DD7777", "scran.knnAnnoy"="#44AA88")
cols <- lapply(pp, FUN=function(x) cols[as.character(sort(unique(x)))])
ha <- HeatmapAnnotation(which = "row", "max\n% lost"=anno_barplot(pclost, bar_width = 1, border=FALSE, gp = gpar(fill = "#282828", col="#282828"), width=unit(1.5,"cm")), df=pp, col=cols, show_legend=FALSE)
legends <- lapply(names(cols), FUN=function(f){
  Legend( labels=names(cols[[f]]), title=f, legend_gp=gpar(fill=cols[[f]]) )
})

do.cluster <- TRUE
H <- NULL
for(i in seq_along(ll2)){
  if((do.cluster || is.null(ro)) && i==1){
    hi <- h2
  }else{
    hi <- FALSE
  }
  if(grepl("silhouette", ll2[[i]]$title)){
    col <- silhscale
  }else{
    col <- pipeComp:::.defaultColorMapping(ll2[[i]]$mat, center=TRUE)
  }
  la <- ra <- NULL
  if(i==length(ll2)) ra <- ha
  name <- ifelse(is.null(ll2[[i]]$name),ll2[[i]]$title,ll2[[i]]$name)
  H <- H + Heatmap(ll2[[i]]$mat, name=name, cluster_rows=hi, row_order=ro2, show_row_names=FALSE, show_column_names=FALSE, row_title=ifelse(i==1,"Unsupervised clustering",""), heatmap_legend_param = list(direction="horizontal", byrow=TRUE, nrow=1), column_title=ll2[[i]]$title, width=unit(ifelse(i<=2,3.2,2.7), "cm"), cluster_columns=ll2[[i]]$cluster_columns, show_column_dend=FALSE, bottom_annotation = pipeComp:::.ds_anno(colnames(ll2[[i]]$mat), legend=FALSE), right_annotation = ra, col=col,  column_names_gp = gpar(fontsize=10), column_title_gp = gpar(fontsize=13), show_heatmap_legend=ifelse(is.null(ll2[[i]]$show_heatmap_legend),TRUE,ll2[[i]]$show_heatmap_legend) )
}

top <- ro2[1:15]
ll3 <- lapply(ll2, FUN=function(x){
  x$mat <- x$mat[top,]
  x
})

ha2 <- HeatmapAnnotation(which = "row", "max\n% lost"=anno_barplot(pclost[top], bar_width = 1, border=FALSE, gp = gpar(fill = "#282828", col="#282828"), width=unit(1.5,"cm")), df=pp[top,], col=cols, show_legend=FALSE)

H2 <- NULL
for(i in seq_along(ll3)){
  if(grepl("silhouette", ll2[[i]]$title)){
    col <- silhscale
  }else{
    col <- pipeComp:::.defaultColorMapping(ll2[[i]]$mat, center=TRUE)
  }
  la <- ra <- NULL
  if(i==length(ll2)) ra <- ha2
  name <- ifelse(is.null(ll2[[i]]$name),ll2[[i]]$title,ll2[[i]]$name)
  H2 <- H2 + Heatmap(ll3[[i]]$mat, name=name, cluster_rows=FALSE, show_row_names=FALSE, row_title="Top 15 analyses", show_column_names=FALSE, heatmap_legend_param = list(direction="horizontal"), column_title=ll2[[i]]$title, width=unit(ifelse(i<=2,3.2,2.7), "cm"), cluster_columns=ll2[[i]]$cluster_columns, show_column_dend=FALSE, bottom_annotation = pipeComp:::.ds_anno(colnames(ll2[[i]]$mat), legend=TRUE), right_annotation = ra, col=col,  column_names_gp = gpar(fontsize=10), column_title_gp = gpar(fontsize=13), show_heatmap_legend=ifelse(is.null(ll2[[i]]$show_heatmap_legend),TRUE,ll2[[i]]$show_heatmap_legend) )
}
```

```{r, fig.width=10, fig.height=7}
pdf("merged.clustered.2.pdf", width=10, height=7)
draw(H, annotation_legend_side="right", heatmap_legend_side="bottom", annotation_legend_list=legends)
dev.off()
draw(H, annotation_legend_side="right", heatmap_legend_side="bottom", annotation_legend_list=legends)
```

```{r, fig.width=10, fig.height=3.5}
pdf("merged.clustered.2.pdf", width=10, height=4)
draw(H2, annotation_legend_side="right", heatmap_legend_side="bottom")
dev.off()
draw(H2, annotation_legend_side="right", heatmap_legend_side="bottom")
```