### 2020-08-04 supplementary_figures/supp_figures.Rmd

---
title: \vspace{2cm}\Large{pipeComp, a general framework for the evaluation of computational pipelines, reveals performant single-cell RNA-seq preprocessing tools}
subtitle: \vspace{2cm}\LARGE{Additional File 1 - Supplementary Figures}\vspace{2cm}
date: "`r format(Sys.time(), '%d %B, %Y')`"
author:
- "Pierre-Luc Germain"
- "Anthony Sonrel"
- "Mark D. Robinson"
geometry: margin=1.5cm
output:
  pdf_document:
    fig_width: 8
    fig_height: 6
    citation_package: natbib
bibliography: ../bmc_article.bib
---

```{r setup, include=FALSE}
# comment this out to print code
knitr::opts_chunk$set(echo=FALSE)
FIG_NB <- 0
getFigNb <- function(increment=FALSE){
  if(increment) FIG_NB <<- FIG_NB + 1
  FIG_NB
}
```

# Fig S`r getFigNb(TRUE)`

```{r, fig.align="center", echo=FALSE, out.height="0.9\\textheight"}
knitr::include_graphics("dataset_description.pdf")
```

### Fig S`r getFigNb()`: Overview of the benchmark datasets

\newpage

```{r child='supp1b_metrics.Rmd', cache=TRUE}
```

```{r child='supp1_dataset_description.Rmd', cache=TRUE}
```

```{r child='supp2_filtering.Rmd', cache=TRUE}
```

```{r child='supp3_normalization.Rmd', cache=TRUE}
```

```{r child='supp4_selection.Rmd', cache=TRUE}
```

```{r child='supp5_dr.Rmd', cache=TRUE}
```

```{r child='supp6_clust.Rmd'}
```

```{r child='supp_overall.Rmd'}
```

```{r child='imputation.Rmd'}
```

```{r child='supp_svadea.Rmd'}
```

# Fig S`r getFigNb(TRUE)`

```{r, fig.align="center", echo=FALSE, out.height="0.8\\textheight"}
knitr::include_graphics("../main_figures/pipeComp_summary.pdf")
```

### Fig S`r getFigNb()`: Summary of the recommendations.