### 2020-08-04 supplementary_figures/supp_svadea.Rmd
updated sva-dea supp

---
title: "Supplementary Figures relative to the SVA-DEA benchmark"
author: "Pierre-Luc Germain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r, include=FALSE}
if(!exists('FIG_NB')){
  FIG_NB <- 0
  getFigNb <- function(increment=FALSE){
    if(increment) FIG_NB <<- FIG_NB + 1
    FIG_NB
  }
}

suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(pipeComp)
  library(grid)
  library(ComplexHeatmap)
})
theme_set(theme_cowplot(font_size = 12))
res12 <- res012 <- res0 <- res1 <- res <- readRDS("../svadea/results/aggregated.rds")
res0$evaluation$dea <- lapply(res$evaluation$dea, FUN=function(x){
  x[which(x$sva.method=="none"),]
})
res1$evaluation$dea <- lapply(res$evaluation$dea, FUN=function(x){
  x[which(x$k==1 & x$sva.method!="none"),]
})
res012$evaluation$dea <- lapply(res$evaluation$dea, FUN=function(x){
  x$k[x$sva.method=="none"] <- NA
  x[which(is.na(x$k) | x$k==2 | x$k==1),]
})
res12$evaluation$dea <- lapply(res$evaluation$dea, FUN=function(x){
  x$k[x$sva.method=="none"] <- NA
  x[which(x$k==2 | x$k==1),]
})
```

# Fig S`r getFigNb(TRUE)`

```{r, fig.width=8.5, fig.height=8.5}
plot_grid(
  dea_evalPlot_curve(res0, agg.by=c("dea.method","filt","minCount"), shapeBy = "dea.method", colourBy = c("filter")),
  dea_evalPlot_curve(res12, agg.by=c("dea.method","filt","minCount"), shapeBy = "dea.method", colourBy = c("filter")),
  dea_evalPlot_curve(res012, agg.by=c("dea.method","sva.method"), shapeBy="dea.method", colourBy = c("sva.method")),
  nrow=3, labels=LETTERS[1:3]
)
```

### Fig S`r getFigNb()`
Accuracy of the differential expression analysis across combinations of: **A:** filters and DEA methods (without any SVA-step), **B:** filters and DEA methods (average across the different SVA strategies using either 1 or 2 surrogate variables), **C:** DEA and SVA methods (using one or two surrogate variables).

\newpage

# Fig S`r getFigNb(TRUE)`

```{r, fig.width=8, fig.height=8.5}
plot_grid(
  grid.grabExpr(draw(
    evalHeatmap(res1, what=c("logFC.pearson", "logFC.mad"), agg.by=c("sva.method","dea.method"), row_split="sva.method", font_factor=.9) +
      evalHeatmap(res1, what=c("TPR", "FDR"), agg.by=c("sva.method","dea.method"), row_split="sva.method", font_factor=.9,
                  filter=threshold==0.05)
  )),
  plotElapsed(res, agg.by=c("sva.method","dea.method")),
  nrow=2, labels=LETTERS[1:2]
)
```

### Fig S`r getFigNb()`
**A:** Accuracy of the estimated logFC (correlation and median absolute deviation from expected logFCs) and of the differential expression analysis (TPR stands for True Positive Rate, and FDR for False Discovery Rate) across the different combinations of SVA and DEA methods (using max 1 dimension). **B:** Running times of the different methods.

\newpage