### 2022-06-27 datasets.Rmd
added code to generate the dataset description

---
title: "datasets overview"
author: "Pierre-Luc Germain"
date: "10/23/2019"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(pipeComp)
})
theme_set(theme_cowplot())
ds <- list.files("datasets/", full.names=TRUE)
names(ds) <- gsub("\\.SCE\\.rds","",basename(ds))

sces <- lapply(ds, FUN=function(x){
  library(scran)
  library(scater)
  sce <- readRDS(x)
  sce <- logNormCounts(sce)
  var.stats <- modelGeneVar(sce)
  sce <- denoisePCA(sce, technical=var.stats)
  reducedDim(sce, "tSNE") <- Rtsne::Rtsne(reducedDim(sce))$Y
  reducedDim(sce, "umap") <- uwot::umap(reducedDim(sce))
  sce
})
```

```{r, fig.width=8.5, fig.height=12}
describeDatasets(sces, pt.size = 0.4, alpha=0.7)
```